---
title: "extract_pdf_table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**tabulizer** provides R bindings to the Tabula java library, which can be used to computationaly extract tables from PDF documents.

- [tabulizer tutorial](https://ropensci.org/tutorials/tabulizer_tutorial.html)
- [Extracting Tables from PDFs in R using the Tabulizer Package | R-bloggers](https://www.r-bloggers.com/extracting-tables-from-pdfs-in-r-using-the-tabulizer-package/)
- [Introducing pdftools - A fast and portable PDF extractor](https://ropensci.org/blog/2016/03/01/pdftools-and-jeroen)

```{r eval=F}
if(!require("ghit")){
    install.packages("ghit")
}

ghit::install_github(c("ropenscilabs/tabulizerjars", "ropenscilabs/tabulizer"))
```


```{r cars}
library(tabulizer)
library(dplyr)

# Location of WARN notice pdf file
pdf = '~/Dropbox/Zotero_storage/5VEVG77A/Berardi - 2013 - Sustainability assessment of urban communities through rating systems.pdf'

# Extract the table
out <- extract_tables(pdf)
```

Got error:

```
Error in .jcall("RJavaTools", "Ljava/lang/Object;", "invokeMethod", cl,  : 
  java.lang.NoSuchMethodError: java.lang.Integer.compare(II)I
```

- [extract_tables() error: java.lang.NoSuchMethodError · Issue #42 · ropensci/tabulizer](https://github.com/ropensci/tabulizer/issues/42#issuecomment-294123978)
- [How do I Update Java for my Mac?](https://www.java.com/en/download/help/mac_java_update.xml)

TODO...:
- [osx - Installed Java 7 on Mac OS X but Terminal is still using version 6 - Stack Overflow](http://stackoverflow.com/questions/12757558/installed-java-7-on-mac-os-x-but-terminal-is-still-using-version-6#answer-33571909)
- [osx - Operation Not Permitted when on root El capitan (rootless disabled) - Stack Overflow](http://stackoverflow.com/questions/32659348/operation-not-permitted-when-on-root-el-capitan-rootless-disabled#answer-32661637)

```bash
sudo ln -s -f "/Library/Internet Plug-Ins/JavaAppletPlugin.plugin/Contents/Home/bin/java" /usr/bin/java
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
